\boldmath Commissioning and performance in beam tests of the highly granular SiW-ECAL technological prototype for the ILC

### 2018-05-11 performance-paper-draft/draft.tex
First draft ready to iterate
Small text and figures improvement

\abstract{
  High precision physics at future colliders as the International Linear Collider (ILC) require unprecedented high precision in the determination
  of the energy of final state particles.
  The needed precision will be achieved thanks to the Particle Flow algorithms (PF) which require highly granular and hermetic calorimeters systems.
  The Silicon-Tungsten Electromagnetic Calorimeter (SiW-ECAL) technological prototype
  design and R\&D is tailored to the baseline design of the ECAL of the International Large Detector (ILD) for the ILC.
  In this document we present the latest news on R\&D of such prototype with fully embedded very front-end (VFE) electronics.
  Special emphasis is given to the presentation and discussion of the commissioning of the prototype
  for beam test and to the performance of the device in such beam test carried at DESY in June 2017.
}

@@

To meet the required precision levels, these detectors will be based on the Particle Flow (PF) techniques\cite{Brient:2002gh,Morgunov:2004ed,Sefkow:2015hna}.
These techniques rely on single particle separation to make possible the choice of the best information available
in the full detector to measure the energy of the final state objects: {\it i.e.} the charged particles momentum is better measured by the tracking system
the photons energy by the electromagnetic calorimeter and the neutral hadrons energy by the full calorimeter system.
In that way the poor resolution of the calorimeter systems (compared with trackers) will  only have a reduced impact in the overall reconstruction.
For this purpose, detectors optimized for PF algorithms have some requirements, summarized here:

@@

The first SiW-ECAL prototype was the so called SiW-ECAL physics prototype.
It was successfully tested at DESY, FNAL and CERN running together with another prototype from the CALICE
collaboration, the analogue hadronic calorimeter AHCAL, delivering the proof of concept of PF calorimetry.
For the physics prototype, the VFE was placed outside the active area with no particular constraints in power consumption.
It consisted of 30 layers of Si as active material alternated with tungsten plates as absorber material.
The active layers were made of a matrix of 3x3 Si wafers. Each of these wafers was segmented in matrices of
6x6 squared pixels of 1x1 $cm^{2}$.
The prototype was divided in 3 modules of 10 layers with different W depth per layer in each of these modules
(0.4, 1.6 and 2.4 $X_{0}$) making a total of 24 $X_{0}$.
Published results proving the good performance of the technology and the PF can be found in references~\cite{Adloff:2011ha,Anduze:2008hq,Adloff:2008aa,Adloff:2010xj,CALICE:2011aa,Bilki:2014uep}.

\section{The SiW-ECAL engineering prototype}

@@

The original design of the silicon wafers included an edge termination made of floating guard-rings.
It was observed in beam test\cite{Cornat:2015eoa,Cornat:2009zz} that the capacitive coupling between such floating guard-rings 
and the pixels at the edge was not negligible in tests with high energy beams (pions and electrons with energies larger than 20-40 GeV).
This coupling lead to fake events in which, at least,
the channels in the four edges of the wafer are triggered at the same time. 
This is why these events are so often called squared events. 
An R\&D program together with Hamamatsu Photonics (HPK Japan) was conducted to study the guard-rings design 
as well as the internal crosstalk. As outcome of this R\&D program, the technological prototype 
is equipped with silicon wafers with:

@@

\begin{figure}[!t]
\centering
\begin{tabular}{l}
  \includegraphics[width=3.0in]{figs/test.jpg} \\
  \includegraphics[width=3.0in]{figs/test.jpg} \\
  \includegraphics[width=3.0in]{figs/test.jpg}
\end{tabular}
\caption{I-V curve, C-V curve, leakage current vs time. \todo{Temporary picture: La Esperanza del Condenado, J. Miró}}
\label{shortslab}
\end{figure}

@@

SKIROC\cite{Callier:2011zz} (Silicon pin Kalorimeter Integrated ReadOut Chip) is the very front end ASIC designed for the readout of the Silicon PIN didoes.
It consists of 64 channels in AMS 0.35 $\mu$m SiGe technology. Each channel comprises a low noise charge preamplifier of variable gain followed by two lines:
a fast shaper for the trigger decision and a set of dual gain slow shaper for charge measurement. 
The gain can be controlled by modifying the feedback capacitance during the configuration of the detector.
With the slowest gain, 6pF, the ASIC will handle a linear dynamic range from 0.1 to up to 1500 MIPs 
(a slightly less than the desired final value for the ILC). 
Finally, a Wilkinson type analogue to digital converter fabricates the digitized charge deposition that can be readout. 
Once one channel is triggered, the ASIC reads out all 64 channels adding a bit of information to tag them as
triggered or not triggered and the information is stored in 15 cell deep physical switched capacitor array (SCA).
The trigger threshold can be controlled both globally and locally (channel by channel) although the
dynamical range for the channel by channel option is too small in the SKIROC version 2 due to a mistake in the development. The trigger holds the voltage of the two slow shapers with a tunable delay and also
holds the voltage swept with the slow clock trigger to obtain timing information.

@@

Every ASU of the SiW-ECAL prototype is equipped with 16 SKIROCs version 2. A new version, 2a, has been produced and will be used to equip new layers currently in production.
In addition, a new generation of the ASIC, SKIROC3, is foreseen for the final detector construction.
In contrast with SKIROC2/2a, the new ASIC will be fully optimized for ILC operaton, {\it i.e.} full zero suppression, reduced power consumption etc.

@@

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
    \includegraphics[width=4in]{figs/ASU_geometry1.eps}  \\
    \includegraphics[width=4in]{figs/ASU_geometry2.eps}  \\
  \end{tabular}
  \caption{Repartition of the ASIC (up) and channels (down) in one ASU. In this perspective, the Si-Sensors are in glued in the back. The channels are separated (in x and y) by 5.5 mm.
    The empty cross in the middle of the ASU corresponds to the 1 mm separation between the sensors.}
\label{ASU}
\end{figure}

The entity
of sensors, thin PCB (printed circuit boards) and ASICs (application-specific integrated circuits) is called Active Signal Units or ASU.
An individual ASU has a lateral dimension of 18x18 cm$^{2}$ and has glued onto it 4 silicon wafers (currently with a thickness of 320 $\mu$m). The high voltage is delivered to the wafers using a HV-kapton sheet that covers the full extension of the wafers.
The ASUs are equipped
further with 16 ASIC for the read out and features 1024 square pads (64 per ASIC) of 5x5 mm. The channels and ASICs are distributed along the ASU as shown in Figure \ref{ASU_geometry}

@@

The design of the subsequent chain of the data acquisition (DAQ)\cite{Gastaldi:2014vaa} system is inspired by the ILC.
Current DAQ consists of three modules which are designed to be generic enough to cope with other applications.
The first module is the so called detector interface (DIF) which is placed at the beginning of each layer holding up to 15 ASUs.
All DIFs are connected by single HDMI cables to the concentrator cards: Gigabit Concentrator Card (GDCC).
The HDMI connection is used to transmit both slow control and data readout.
One GDCC controls up to 7 DIFs collecting all data from them and distributing to them the system clock and fast commands.
The most downstream module is the clock and control card (CCC).
The CCC provides a clock, control fan-out of up to 8 GDCCs and accepts and distributes external signals (i.e. spill signals).

@@

The prototype tested in beam is composed of 7 layers consisting in one short SLAB each.
A short SLAB is made of: one single ASU FEV11 equipped with 16 SKIROC2 and 4 wafers glued on the other side (1024 channels);
an interface card SMBv4; and a DIF. All these elements sit on a "U" shape carbon structure to protect the wafers.
The full system is then covered by two aluminum plates to provide electromagnetic shielding and mechanical stability.
The 7 layers are supported by a PVC and aluminum structure that in which can be inserted to 10 layers.
The distance between two layer slots is 15 mm and this space is used to place tungsten plates of different thicknesses.
For the beam test described in Section \ref{sec:beamtest}, the layers were placed in positions 1-6 and 10.
During the commissioning, the slabs have been tested both inside and outside of the structure.
A photograph showing the prototype can be seen in Figure \ref{proto}. The ASICs and channels
are labeled with numbers following design and DAQ criteria: from 0-16 in the case of the ASICs and from 0-63 in the case of the channels.
%The mapping from this numbering scheme to real positions in the ASUs is shown in the right plot of Figure \ref{proto}.

@@

The SKIROC2 has been already commissioned for previous experiences (see for example the 
references \cite{Amjad:2014tha,Suehara:2018mqk}). 
Therefore, all the settings used for the SKIROC2 to obtain 
the results described in this paper are motivated by previous experiences. 
For example, the value of 1.2pF for the preamplifier. 
With this gain, the SKIROC2 ensures a linearity better than 90\% 
for 0.5-200 MIPs, which is optimal for 
electromagnetic showers created by few GeV 
electrons or positrons. The delay of the trigger signal is the same than in 
reference \cite{Amjad:2014tha,Suehara:2018mqk}: 130 DAC for all ASICS. 
This value was chosen by selecting the delay in which an injected signal
of few MIPs of value get the maximum value of ADC. %Finally, the value of the
%compensation capacitance at the output of the preamplifier has been set to 6pF instead of 
%4pF to prevent overshots and ripples of the signal used as input to the trigger
%decision and ADC lines.

@@

The purpose of this phase was to monitor and control the
noise levels of the prototype. The commissioning includes the definition of trigger threshold values and a list of noisy channels to be masked.
Studying and control the noise levels in an self-triggered and high-granularity calorimeter
is crucial since noisy channels may saturate the DAQ faster than physical signals.
Two different types of noise were identified:
randomly distributed (in time) noisy channels 
and noise bursts affecting to all layers at the same time.

The first type of noise source (randomly distributed in time)
forced us to define a list of channels to be masked in all 
layers. Again, two different type of noisy channels have been identified. 
The first group is compose by randomly distributed (in space) channels which 
slightly larger number of triggers than the others. The second group is made of
a fixed list of channels, all in the same positions for all SLABs, that were giving
underflowed ADC triggers at high rates.

The channels associated to "underflowed-noise" events were identified by means of the online monitoring
tools. Some preliminary inspections of the FEV11 layout hint that
the cause of this "underflowed-noise" events may be in an improvable routing of the
lines from the PIN pad to the SKIROC. 
For this reason, if a given channel appeared to be associated 
to "underflowed-noise" events in one of the SLABs, 
it was masked in all of them, taking in this way the most conservative solution.
For example, all channels number 37 were masked in all ASICs.
More studies in this respect are to come.

The list of the noisy channels randomly distributed in time and space was
defined by means of dedicated data taking runs. 
These runs were characterized by: their short acquisition windows (of 2.5 ms, only 1.1 ms 
of effectively data taking time), their low
spill frequency (5 Hz) to minimize the chances of having real events due to cosmic rays 
hitting the detector during the
data taking; and the relatively high trigger threshold values between 250 and 400 DAC 
(which is equivalent to $\sim$0.5-2 MIP, see Section \ref{sec:comm_trigger} for more 
information). The full process was an iterative process starting with several repetitions
of runs at a high threshold value and then some iterations at lower threshold each up to 250 DAC. 
In every of there iterations, the channels identified as noisy were masked.
A full run involved 3-5 threshold values and at least two repetitions per value.
Again, we decided to follow a conservative approach for list the noisy channels to be masked:
if the channel was triggered at rates larger than 0.5-1\% of the total number of triggers it was added to the list.

@@

We observed that the noise bursts happening coherently at the same time in all SLABS
was associated to cases when the electrical isolation between SLABs was broken.
This is a hint of a system effect as grounding loops or noise in the power supplies.
The issue was circumvented by improving the electrical isolation of single layers.
We have also observed that the noise burst happen only at the end of long acquisitions therefore, in addition to the improved isolation,
we selected short enough acquisitions windows (which indeed are the most appropriate to the high rates of particles in the DESY beam).
However, more studies in the laboratory are needed in order to fully understand these issues.

@@

Once that the SKIROC main settings have been fixed and the masking of the noisy channels at a relatively high threshold value
has been done, the final step is the selection of the optimal trigger threshold values.
For that, we perform dedicated runs in which we scan a reasonable range of trigger threshold values (in DAC units)
with all channels enabled (excepted the marked as noisy). For each of the DAC points
we took data during $\sim$ 1 minute per DAC in order to speed up the commissioning.
Plotting the total number of hits normalized to the maximum vs the DAC
for each channel gives the so-called S-curves. In the absence of external signals (cosmic rays, injected signals, etc) 
these noise S-curves show the convolution of the envelope of the 
electronic noise at the output of the fast shaper (the trigger decision line).
Indeed, the size of this enveloppe is related to slow clock frequency
(larger periods, larger possibility to find a noise variation over the threshold).

To reduce to the minimum the
presence of cosmic rays signals, we perform runs with short open acquisition windows (1.1 ms). 
In Figure \ref{scurve_channels} 
two S-curves of two different channels from the second layer of the setup are shown.

@@

\begin{equation}
DAC_{optimal}^{ASIC-j} = <p_{1}^{ASIC-j}> + 5 \times <p_{2}^{ASIC-j}>
\end{equation}
and 230 if more than the 30\% of the 64 channels S-curves in the ASIC could be fitted. 
If not, we took a most conservative value of 250 DAC. The election of these two values was 
sustained by previous experiences, see for example \cite{Amjad:2014tha}.

@@

In Figure \ref{scurves_injection} we see the 1 MIP and 2 MIP S-curves obtained for a one channel
in a SKIROC testboard in which a single SKIROC2 in BGA package is placed and the 1 MIP and 2 MIPs 
"fake" signals are directly injected in the preamplifier (via a 3 pF capacitor in the injection line). 
From this plot we can extract a S/N ratio of $\sim$12. 
We don expect large differences with the results that we would obtain with 
full equipped SLABs 
although this board is thought for commissioning and test of the SKIROC ASICS in an
"ideal" environment in contrast with the FEV
ASUS that are optimized to meet detector constraints and to hold several ASICs at the same time.
For example: only power pulsing of the analog power is done in this testboard,
there is a set of external capacitors on the board connected to some ASIC pins and
usually only one channel is tested. This allows us to commission the SKIROC2 itself

@@

We have obtained similar results using real signals, in this case cosmic rays signals. This is shown 
in Figure \ref{scurves_cosmics}  where 
we show the result of the fit to the noise S-curves for all channels (in red) in one of the ASICs of 
the second layer together with the results of the S-curve with cosmic rays (black points and blue 
line). For the cosmic S-curve we took 30 min of data per point using very long spills (150 ms) in 
frequencies of 5Hz and we add up the triggers from all channels together. In addition, using the 
knowledge of the MIP value (see next section), we performed a simple cut in the ADC
to keep only signals corresponding to a MIP 
(pedestal subtracted) $\pm30$ ADC counts. With this cut we try reduce the contribution of muons 
traversing the wafers with very large angles but we don't have a way to select perpendicular tracks 
since we were using simple setups with only single slabs during in the run. For all these reasons, we expect
a broader distribution of the cosmic S-curve. This is one of the reasons why, if we calculate the S/N 
using this plot, the value would be unrealistic (much lower, indeed). 
In addition to this, we should remember that the noise S-curve do correspond to the real noise 
distribution but only to the envelope of the noise in the fast shaper, therefore the
distance between the two S-curves is smaller than the real distance between noise and signal.

@@

In conclusion, both ways of estimating the S/N ratio of the trigger have their own limitations but in 
both cases, the 50\% of the S-curve of 1 MIP are compatible within $\pm10$ ADC. This allow us
to, at least, estimate the relative distance between our final trigger threshold for every ASIC and 
the 1 MIP position. This is shown in Figure \ref{trigger_thresholds}. Two ASICS ($\sim2\%$) have a
trigger threshold value of 0.67 MIP. The 11\% of the ASICs have a trigger threshold between 0.5 and 0.55 MIP and the rest, 87\%, have trigger thresholds slightly lower 0.5 MIP.

\begin{figure}[!t]
  \centering
  \includegraphics[width=4in]{figs/commissioning/threshold_chip.eps}
  \caption{Summary of trigger threshold settings for all ASICs in the setup. The values are given in DAC and in estimated MIP units, using the distance between two MIPs from Figure \ref{scurves_injection}. The uncertainty quoted in the right axis is estimated from difference between the 50\% position of the S-curve for one MIP extracted from this figure and Figure \ref{scurves_cosmics}.}
\label{trigger_thresholds}
\end{figure}

@@

All the commissioning procedure described above relies in very conservative decisions
due to the presence of unknown noise sources during most of the commissioning phase. These sources are
now well know and isolated and therefore a new ``noise'' commissioning procedure can proposed
for the next beam test. It will consist in an iterative algorithm that first
will identify and mask the "underflowed-noisy" channels and afterwards run a set of acquisitions
in which the number of triggers per channel will be compared with the number of expected triggers
assuming only cosmics rays as signal. This will allow us to have an
unambiguous definition of the noise levels channel per channel instead of 
defining that levels relatively to the total number
of recorded triggers. Finally, once that the noisy channels
are identified, the optimization of the threshold levels
will be performed and the last run for identification of noisy channels will be taken using this optimal
level will be taken. In this run, we can decide to mask the noisy channels or
to increase its own trigger threshold level if we are using Skiroc2a.

Using this new procedure we manage to reduce the number of masked channels by a factor 2 without any loss of performance,
at least in the laboratory and using 3 of the 7 SLABs. This new procedure will be tested in the next beam test.
Also, in order to
optimize the commissioning of the detector,
we will propose a new set of measurements in the next beam test consisting in:
a scan of optimal delay of the hold values of the trigger using MIP like particles;
and a threshold scan for the determination of the S/N in the trigger line using
1 MIP and $\sqrt{2}$ MIP (tilting the detector by 45 degrees) signals.

@@

\subsubsection{S/N ratio in the ADC}
\label{sec:mip}

>>>>

\abstract{
  High precision physics at future colliders as the International Linear Collider (ILC) require unprecedented high precision in the determination
  of the energy of final state particles.
  The needed precision will be achieved thanks to the Particle Flow algorithms (PF) which require compact, highly granular and hermetic calorimeters systems.
  The Silicon-Tungsten Electromagnetic Calorimeter (SiW-ECAL) technological prototype
  design and R\&D is tailored to the baseline design of the ECAL of the International Large Detector (ILD) for the ILC.
  In this document we present the latest news on R\&D of such prototype with fully embedded very front-end (VFE) electronics.
  Special emphasis is given to the presentation and discussion of the commissioning of the prototype
  for beam test and to the performance of the device in such beam test carried at DESY in June 2017.
}

@@

It is known that in a {\it classical} typical multipurpose detector
of a collider experiment the charged particles momentum is better measured by the tracking system
the photons energy are only observed in the calorimeter system (mainly in the electromagnetic) if they are not converted
and the measurement of the neutral hadrons energy can only be done by involving the full calorimeter systems.
To meet the required precision levels by the ILC or CLIC physics goals,
new techniques relying on single particle separation to make possible the choice of the best information available
in the full detector to measure the energy of the final state objects have been developed.
These techniques are called Particle Flow (PF) techniques \cite{Brient:2002gh,Morgunov:2004ed,Sefkow:2015hna}
and allow to reduce the impact of the poor resolution of the calorimeter systems (compared with trackers) in the overall reconstruction.
For this purpose, detectors optimized for PF algorithms have some requirements. Some of them are summarized here:

@@

The first SiW-ECAL prototype was the so called SiW-ECAL physics prototype.
It was successfully tested at DESY, FNAL and CERN running in front of another prototype from the CALICE
collaboration, the analogue hadronic calorimeter AHCAL, delivering the proof of concept of PF calorimetry.
For the physics prototype, the VFE was placed outside the active area with no particular constraints in power consumption.
It consisted of 30 layers of Si as active material alternated with tungsten plates as absorber material.
The active layers were made of a matrix of 3x3 Si wafers. Each of these wafers was segmented in matrices of
6x6 squared pixels of 1x1 $cm^{2}$.
The prototype was divided in 3 modules of 10 layers with different W depth per layer in each of these modules
(0.4, 1.6 and 2.4 $X_{0}$) making a total of 24 $X_{0}$.
Published results proving the good performance of the technology and the PF can be found in references~\cite{Adloff:2011ha,Anduze:2008hq,Adloff:2008aa,Adloff:2010xj,CALICE:2011aa,Bilki:2014uep}.

\section{The SiW-ECAL technological prototype}

@@

The original design of the silicon wafers included an edge termination made of floating guard-rings.
It was observed in beam tests \cite{Cornat:2015eoa,Cornat:2009zz} that the capacitive coupling between such floating guard-rings 
and the pixels at the edge was not negligible in tests with high energy beams (pions and electrons with energies larger than 20-40 GeV).
This coupling lead to fake events in which, at least,
the channels in the four edges of the wafer are triggered at the same time. 
This is why these events are so often called squared events. 
An R\&D program together with Hamamatsu Photonics (HPK Japan) was conducted to study the guard-rings design 
as well as the internal crosstalk. As outcome of this R\&D program, the technological prototype 
is equipped with silicon wafers with:

@@

\begin{figure}[!t]
\centering
\begin{tabular}{lll}
  \includegraphics[width=1.8in]{figs/test.jpg} &
  \includegraphics[width=1.8in]{figs/test2.jpg} &
  \includegraphics[width=1.8in]{figs/test3.jpg}
\end{tabular}
\caption{I-V curve, C-V curve, leakage current vs time. \todo{Temporary picture: Triptic La Esperanza del Condenado, J. Miró}}
\label{shortslab}
\end{figure}

@@

SKIROC\cite{Callier:2011zz} (Silicon pin Kalorimeter Integrated ReadOut Chip) is the very front end ASIC (application-specific integrated circuits)
designed for the readout of the Silicon PIN didoes.
In its version SKIROC2 (and 2a) it consists of 64 channels in AMS 0.35 $\mu$m SiGe technology.
Each channel comprises a low noise charge preamplifier of variable gain followed by two lines:
a fast shaper for the trigger decision and a set of dual gain slow shaper for charge measurement. 
The gain can be controlled by modifying the feedback capacitance during the configuration of the detector.
With the slowest gain, 6pF, the ASIC will handle a linear dynamic range from 0.1 to up to 1500 MIPs 
(a slightly less than the desired final value for the ILC). 
Finally, a Wilkinson type analogue to digital converter fabricates the digitized charge deposition that can be readout. 
Once one channel is triggered, the ASIC reads out all 64 channels adding a bit of information to tag them as
triggered or not triggered and the information is stored in 15 cell deep physical switched capacitor array (SCA).
The trigger threshold can be controlled both globally and locally (channel by channel) although the
dynamical range for the channel by channel option is too small in the SKIROC version 2 due to a mistake in the development. This issue has
been solved in the SKIROC2a. The trigger holds the voltage of the two slow shapers with a tunable delay and also
holds the voltage swept with the slow clock trigger to obtain timing information.

@@

Every ASU of the SiW-ECAL prototype is equipped with 16 SKIROCs version 2. A newer version, 2a, with some incremental improvements
and bug fixing has been produced and will be used to equip new layers currently in production.
In addition, a new generation of the ASIC, SKIROC3, is foreseen for the final detector construction.
In contrast with SKIROC2/2a, the new ASIC will be fully optimized for ILC operaton, {\it i.e.} full zero suppression, reduced power consumption etc.

@@

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
    \includegraphics[width=4in]{figs/ASU_geometry1.eps}  \\
    \includegraphics[width=4in]{figs/ASU_geometry2.eps}  \\
  \end{tabular}
  \caption{Repartition of the ASIC (up) and channels (down) in one ASU. In this perspective, the Si-Sensors are in glued in the back.
    The channels are separated (in x and y) by 5.5 mm.
    The empty cross in the middle of the ASU corresponds to the 1 mm separation between the sensors.
    The ASICs and channels
    are labeled with numbers following design and DAQ criteria: from 0-16 in the case of the ASICs and from 0-63 in the case of the channels.
  }
\label{ASU}
\end{figure}

The entity
of sensors, thin PCB (printed circuit boards) and ASICs is called Active Signal Units or ASU.
An individual ASU has a lateral dimension of 18x18 cm$^{2}$ and has glued onto it 4 silicon wafers (currently with a thickness of 320 $\mu$m). The high voltage is delivered to the wafers using a HV-kapton sheet that covers the full extension of the wafers.
The ASUs are equipped
further with 16 ASIC for the read out and features 1024 square pads (64 per ASIC) of 5x5 mm.
The channels and ASICs are distributed along the ASU as shown in Figure \ref{ASU}

@@

The design of the subsequent chain of the data acquisition (DAQ)\cite{Gastaldi:2014vaa} system is inspired by the ILC.
Current DAQ consists of three modules which are designed to be generic enough to cope with other applications.
The first module is the so called detector interface (DIF) which is placed at the beginning of each layer holding up to 15 ASUs.
All DIFs are connected by single HDMI cables to the concentrator cards: Gigabit Concentrator Card (GDCC).
The HDMI connection is used to transmit both slow control and data readout.
One GDCC controls up to 7 DIFs collecting all data from them and distributing among them the system clock and fast commands.
The most downstream module is the clock and control card (CCC).
The CCC provides a clock, control fan-out of up to 8 GDCCs and accepts and distributes external signals (i.e. spill signals).

@@

The prototype tested in beam is composed of 7 layers consisting in one short SLAB each.
A short SLAB is made of: one single ASU FEV11 equipped with 16 SKIROC2 and 4 wafers glued on the other side (1024 channels);
an interface card SMBv4; and a DIF. All these elements sit on a "U" shape carbon structure to protect the wafers.
The full system is then covered by two aluminum plates to provide electromagnetic shielding and mechanical stability.
The 7 layers are supported by a PVC and aluminum structure that in which can be inserted to 10 layers.
The distance between two layer slots is 15 mm and this space is used to place tungsten plates of different thicknesses.
For the beam test described in Section \ref{sec:beamtest}, the layers were placed in positions 1-6 and 10.
During the commissioning, the slabs have been tested both inside and outside of the structure.
A photograph showing the prototype can be seen in Figure \ref{proto}. 
%The mapping from this numbering scheme to real positions in the ASUs is shown in the right plot of Figure \ref{proto}.

@@

The SKIROC2 has been already commissioned for previous experiences (see for example the 
references \cite{Amjad:2014tha,Suehara:2018mqk}). 
Therefore, all the settings used for the SKIROC2 to obtain 
the results described in this paper are motivated by these previous experiences. 
For example, the value of 1.2pF for the preamplifier. 
With this gain, the SKIROC2 ensures a linearity better than 90\% 
for 0.5-200 MIPs\footnote{\todo{Is this a fact?}}, which is enough for 
electromagnetic showers created by few GeV 
electrons or positrons. The delay of the trigger signal is the same than in 
reference \cite{Amjad:2014tha,Suehara:2018mqk}: 130 DAC for all ASICS. 
This value was chosen by selecting the delay in which an injected signal
of few MIPs of value get the maximum value of ADC. %Finally, the value of the
%compensation capacitance at the output of the preamplifier has been set to 6pF instead of 
%4pF to prevent overshots and ripples of the signal used as input to the trigger
%decision and ADC lines.

@@

The purpose of this phase was to monitor and control the
noise levels of the prototype. The commissioning includes the definition of trigger threshold values and a list of noisy channels to be masked.
Studying and control the noise levels
is crucial since noisy channels may saturate the DAQ faster than physical signals.
Two different types of noise were identified:
randomly distributed in time noisy channels 
and noise bursts affecting to all layers at the same time.

The first type of noise source (randomly distributed in time)
forced us to define a list of channels to be masked in all 
layers. Again, two different type of noisy channels have been identified. 
The first group is compose by randomly distributed in space channels.
The second group is made of
a fixed list of channels, all in the same positions for all SLABs, that were giving
underflowed ADC triggers at high rates. 
The channels associated to "underflowed-noise" events were identified by means of the online monitoring
tools. Some preliminary inspections of the FEV11 layout hint that
the cause of this "underflowed-noise" events may be in an improvable routing of the
lines from the PIN pad to the SKIROC. 
For this reason, if a given channel appeared to be associated 
to "underflowed-noise" events in one of the SLABs, 
it was masked in all of them, taking in this way the most conservative solution.
For example, all channels number 37 were masked in all ASICs.
More studies in this respect are to come.

The list of the noisy channels randomly distributed in time and space was
defined by means of dedicated data taking runs. 
These runs were characterized by: their short acquisition windows (of 2.5 ms, only 1.1 ms 
of effectively data taking time), their low
spill frequency (5 Hz) to minimize the chances of having real events due to cosmic rays 
hitting the detector during the
data taking; and the relatively high trigger threshold values between 250 and 400 DAC 
(which is equivalent to $\sim$0.5-2 MIP, see Section \ref{sec:comm_trigger} for more 
information). The full process was an iterative process starting with several repetitions
of runs at a high threshold value and then some iterations at lower threshold each up to 250 DAC. 
In every of there iterations, the channels identified as noisy were masked.
A full run involved 3-5 threshold values and at least two repetitions per value.
Again, we decided to follow a conservative approach for list the noisy channels to be masked:
if the channel was triggered at rates larger than 0.5-1\% of the total number of triggers per ASIC
it was added to the list.

@@

We observed that the noise bursts happening coherently at the same time in all SLABS
was associated to cases when the electrical isolation between SLABs was broken.
This is a hint of a system effect as can be the appearence of
grounding loops or noise in the power supplies.
The issue was circumvented by improving the electrical isolation of single layers.
We have also observed that the noise bursts happened only at the end of long acquisitions therefore, in addition to the improved isolation,
we selected short enough acquisitions windows (which indeed are the most appropriate to the high rates of particles in the DESY beam).
However, more studies in the laboratory are needed in order to fully understand these issues.

@@

Once that the SKIROC main settings have been fixed and the masking of the noisy channels at a relatively high threshold value
has been done, the final step is the selection of the optimal trigger threshold values.
For that, we perform dedicated runs in which we scan a reasonable range of trigger threshold values (in DAC units)
with all channels enabled (excepted the marked as noisy). For each of the DAC points
we took data only during $\sim$ 1 minute per DAC in order to speed up the commissioning.
Plotting the total number of hits normalized to the maximum vs the DAC
for each channel gives the so-called S-curves. In the absence of external signals (cosmic rays, injected signals, etc) 
these noise S-curves show the convolution of the envelope of the 
electronic noise at the output of the fast shaper (the trigger decision line).
Indeed, the size of this enveloppe is related to the slow clock frequency
(larger periods increase the possibility to find a noise variation over the threshold).

To reduce to the minimum the
presence of cosmic rays signals, we perform runs with short open acquisition windows (1.1 ms)\footnote{The acquisition window is defined by means
  of a external pulse generator that emulates similar spill conditions than at the ILC. For DAQ needs, the actual data recording starts 1.4 ms after
rising edge of the spill signal arrives. In this paper we will always refer to the effectively open acquisition window not including the mentioned 1.4 ms.}. 
In Figure \ref{scurve_channels} 
two S-curves of two different channels from the second layer of the setup are shown.

@@

\begin{equation}
DAC_{optimal}^{ASIC-j} = <p_{1}^{ASIC-j}> + 5 \times <p_{2}^{ASIC-j}>
\end{equation}
and 230 if at least the the 30\% of the 64 channels S-curves in the ASIC could be fitted. 
If less than the 30\% of the channels were fit
we took a most conservative value of 250 DAC for the threshold of the ASIC.
The election of these two values was 
sustained by previous experiences, see for example \cite{Amjad:2014tha}.

@@

In Figure \ref{scurves_injection} we see the 1 MIP and 2 MIP S-curves obtained for one channel
in a SKIROC testboard in which a single SKIROC2 in BGA package is placed and the 1 MIP and 2 MIPs 
"fake" signals are directly injected in the preamplifier (via a 3 pF capacitor located in the injection line as shown in Figure \ref{SKIROC2}). 
From this plot we can extract a S/N ratio of $\sim$12. 
We don expect large differences with the results that we would obtain with 
full equipped SLABs 
although this board is thought for commissioning and test of the SKIROC ASICS in an
"ideal" environment in contrast with the FEV
ASUS that are optimized to meet detector constraints and to hold several ASICs at the same time.
For example: only power pulsing of the analog power is done in this testboard,
there is a set of external capacitors on the board connected to some ASIC pins and
usually only one channel is tested. This board allows us to commission and debug the SKIROC itself
but not the full system.

@@

We have obtained similar results using real signals, in this case cosmic rays signals. This is shown 
in Figure \ref{scurves_cosmics}  where 
we show the result of the fit to the noise S-curves for all channels (in red) in one of the ASICs of 
the second layer together with the results of the S-curve with cosmic rays (black points and blue 
line). For the cosmic S-curve we took 30 min of data per point using very long spills (150 ms) in 
frequencies of 5Hz and we add up the triggers from all channels together. In addition, using the 
knowledge of the MIP value (see next section), we performed a simple cut in the ADC
to keep only signals corresponding to a MIP 
(pedestal subtracted) $\pm30$ ADC counts. With this cut we try reduce the contribution of muons 
traversing the wafers with very large angles but we don't have a way to select perpendicular tracks 
since we were using simple setups with only single slabs. For all these reasons, we expect
a broader distribution of the cosmic S-curve. This is one of the reasons why, if we calculate the S/N 
using this plot, the value would be unrealistic (much lower, indeed). 
In addition to this, we should remember that the noise S-curve do correspond to the real noise 
distribution but only to the envelope of the noise in the fast shaper, therefore the
distance between the two S-curves is smaller than the real distance between noise and signal.

@@

In conclusion, both ways of estimating the S/N ratio of the trigger have their own limitations but in 
both cases, the 50\% of the S-curve of 1 MIP are compatible within $\pm10$ ADC. This allow us
to, at least, estimate the relative distance between our final trigger threshold for every ASIC and 
the 1 MIP position by combining the information extracted from Figs \ref{scurves_injection} and \ref{scurves_cosmics}.
This is shown in Figure \ref{trigger_thresholds}. Two ASICS ($\sim2\%$) have a
trigger threshold value of 0.67 MIP. The 11\% of the ASICs have a trigger threshold between 0.5 and 0.55 MIP and the rest, 87\%, have trigger thresholds slightly lower 0.5 MIP.

\begin{figure}[!t]
  \centering
  \includegraphics[width=4in]{figs/commissioning/threshold_chip.eps}
  \caption{Summary of trigger threshold settings for all ASICs in the setup. The values are given in DAC and in estimated MIP units, using the distance between two MIPs from Figure \ref{scurves_injection}. The uncertainty quoted in the right axis is estimated from difference between the 50\% position of the S-curve for one MIP extracted from Figured \ref{scurves_injection} and \ref{scurves_cosmics}.}
\label{trigger_thresholds}
\end{figure}

@@

All the commissioning procedure described above relies in very conservative decisions
due to the presence of unknown noise sources during most of the commissioning phase. These sources are
now well know and isolated and therefore a new ``noise'' commissioning procedure has been studied.
It will consist in an iterative algorithm that first
will identify and mask the "underflowed-noisy" channels and afterwards run a set of acquisitions
in which the number of triggers per channel will be compared with the number of expected triggers
assuming only cosmics rays as signal. This will allow us to have an
unambiguous definition of the noise levels channel per channel instead of 
defining such levels relatively to the total number
of recorded triggers per ASIC. Finally, once that the noisy channels
are identified, the optimization of the threshold levels
will be performed and a last run for identification of noisy channels will be taken using this optimal
threshold. In this run, we can decide to mask the noisy channels or
to increase its own trigger threshold level if we are using Skiroc2a.

Using this new procedure we manage to reduce the number of masked channels by a factor 2 without any loss of performance,
at least in the laboratory and using 3 of the 7 SLABs. This new procedure will be tested in the next beam test.
Also, in order to
optimize the commissioning of the detector,
we propose a new set of measurements in the next beam test such as
a scan of optimal delay of the hold values of the trigger using MIP like particles
and a threshold scan for the determination of the S/N in the trigger line. This later
can be done comparing 1 MIP and $\sqrt{2}$ MIP (tilting the detector by 45 degrees) signals.

@@

\subsubsection{S/N ratio in the ADC}
\label{sec:sn}
